---
title: "ForeCasting Methods Project"
author: "Group"
date: "01/05/2022"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
#loading libraries
library(fpp3)
library(tidyverse)
#working directory for github
setwd("~/Project-Forecasting-Methods")

#--------------------------------------------- be careful with this code below

#loading the data -->just to run for the first time 

#GlobalTemperature <-read.csv("data/GlobalTemperatures.csv")

#adding column -->just to run for the first time

#GlobalTemperature["Country"]<- "Global"

#creating new csv file --> just to run for the first time

#write.csv(GlobalTemperature, "GlobalTemperatures2.csv")

#Country <- read.csv("data/GlobalLandTemperaturesByCountry.csv")
#names(Country) <-c("dt","LandAverageTemperature", "AverageTemperatureUncertainty","Country")
#write.csv(Country,"GlobalLandTemperaturesByCountry2.csv")

#------------------------------------------- be careful with this code above



```


```{r, message=FALSE, include=FALSE}
#merging Global file with country file
GlobalCountryTemp <- list.files(pattern = "*.csv", full.names = TRUE) %>% 
  lapply(read_csv) %>%                             
  bind_rows  

#transforming in a tsibble

GlobalCountryTemp<- GlobalCountryTemp%>% mutate(Date = yearmonth(dt)) %>% as_tsibble(index=Date, key="Country") %>% select(Date, Country, everything()) %>% select(-dt,-LandMaxTemperatureUncertainty, -LandMinTemperatureUncertainty,-AverageTemperatureUncertainty) 

#changing columns names
names(GlobalCountryTemp) <-c('Date','Country','idk', 'LandAvgTemp','LandAvgTempUnc', 'LandMaxTemp', 'LandMinTemp', 'LandOceanAvgTemp', 'LandOceanAvgTempUnc')
GlobalCountryTemp <-GlobalCountryTemp %>%  select(-idk)
#-------------------Objects for visualization

#Creating GlobalTemp object, which contains only Global values and cleaning it for visualization
GlobalTempV <- GlobalCountryTemp %>% filter(Country=="Global")  %>% filter(year(Date)>1984, year(Date)<2014)
GlobalTempV <-na.omit(GlobalTempV)

#Creating CountryTemp objects, each contains one country and cleaning it for visualizations purpose -->Not sure if its needed to use
#using years above 1984 for visualisation purposes

#PortugalTempV <-GlobalCountryTemp %>% filter(Country == "Portugal") %>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)
#GermanyTempV <-GlobalCountryTemp %>% filter(Country == "Germany")%>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)
#UkTempV <-GlobalCountryTemp %>% filter(Country == "United Kingdom")%>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)
#BrazilTempV <-GlobalCountryTemp %>% filter(Country == "Brazil")%>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)
#USATempV <-GlobalCountryTemp %>% filter(Country == "United States")%>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)
#ChinaTempV <-GlobalCountryTemp %>% filter(Country == "China")%>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)
#MaliTempV <-GlobalCountryTemp %>% filter(Country == "Mali")%>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)
#SouthAfricaTempV <-GlobalCountryTemp %>% filter(Country == "South Africa")%>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)
#AustraliaTempV <-GlobalCountryTemp %>% filter(Country == "Australia")%>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)
#GreenlandTempV <- GlobalCountryTemp %>% filter(Country == "Greenland")%>% select(Date, Country, LandAvgTemp)%>% filter(year(Date)>1984)


#divided it because analysing 10 at once gets really messy
countries<- c('Portugal','Mali','Greenland')
countries2<- c('Portugal', 'Germany', 'United Kingdom','Brazil','United States')
countries3<-c('China','Mali','South Africa','Australia','Greenland')

#-------------------Objects to use in the rest of the work

#Global
GlobalTemp <- GlobalCountryTemp %>% filter(Country=="Global") 
GlobalTemp <-na.omit(GlobalTemp)

#Countries-->not sure if its needed to use
PortugalTemp <-GlobalCountryTemp %>% filter(Country == "Portugal") %>% select(Date, Country, LandAvgTemp)
GermanyTemp <-GlobalCountryTemp %>% filter(Country == "Germany")%>% select(Date, Country, LandAvgTemp)
UkTemp <-GlobalCountryTemp %>% filter(Country == "United Kingdom")%>% select(Date, Country, LandAvgTemp)
BrazilTemp <-GlobalCountryTemp %>% filter(Country == "Brazil")%>% select(Date, Country, LandAvgTemp)
USATemp <-GlobalCountryTemp %>% filter(Country == "United States")%>% select(Date, Country, LandAvgTemp)
ChinaTemp <-GlobalCountryTemp %>% filter(Country == "China")%>% select(Date, Country, LandAvgTemp)
MaliTemp <-GlobalCountryTemp %>% filter(Country == "Mali")%>% select(Date, Country, LandAvgTemp)
SouthAfricaTemp <-GlobalCountryTemp %>% filter(Country == "South Africa")%>% select(Date, Country, LandAvgTemp)
AustraliaTemp <-GlobalCountryTemp %>% filter(Country == "Australia")%>% select(Date, Country, LandAvgTemp)
GreenlandTemp <- GlobalCountryTemp %>% filter(Country == "Greenland")%>% select(Date, Country, LandAvgTemp)
```
## 2 Time series graphics

  In this section the group will explore the dataset by ploting time series graphics and analyzing them, this section is really important as it provides the group an overall view of the temperature changing through the time and allows the group to identify possible patterns, such as trends, cycles and seasonalities during the time. 

### General overview of the dataset

```{r, warning = FALSE}
GlobalTemp %>% autoplot(LandOceanAvgTemp)+
  geom_line(color="blue")+
  labs(title = "Land & Ocean Temperature - Global",
       subtitle = "Average",
       y="LandOceanAvgTemp Cº",
       caption = "fig. 1")

GlobalCountryTemp %>% filter(Country == "Brazil") %>% autoplot(LandAvgTemp)+
  geom_line(color="green")+
  labs(title = "Land Temperature in Brazil - Since 1850",
       subtitle="Average",
       y="LandAvgTemp Cº",
       caption = "fig. 2")

GlobalCountryTemp %>% filter(Country=="Global") %>% autoplot(LandAvgTempUnc)+
  labs(title = "Land Average temperature Uncertainty - Evolution through time",
       caption = "fig. 3")
```

  The dataset provides three different views that can be combined and analysed in multiple different ways.

  The first view is the Global Land & Ocean average temperature (see fig. 1), it provides an interesting sight on how the combination of both land and ocean temperature has been changing since 1850.

  The central view of the dataset is the Land temperature analysis, which can be divided into two parts: 

* A global analysis that contains average, max and min temperature values 
* A more narrow analysis by country that allows us to observe particular countries average temperatures over the years. Note that the years in analyse may differ from country to country due to missing data in the dataset, especially for years before 1900. Figure 2 shows the time series of Land Temperature of Brazil.


The final view is related with the uncertainty of the average temperature values and how it has changed over time (see fig. 3).

```{r, warning=FALSE}
GlobalTempV %>% ggplot()+
  geom_line(data=GlobalTempV, aes(Date, LandMaxTemp, col="LandMaxTemp"))+
  geom_line(data=GlobalTempV, aes(Date,LandAvgTemp,col="LandAvgTemp"),lwd=2)+
  geom_line(data=GlobalTempV, aes(Date, LandMinTemp, col="LandMinTemp"))+
  labs(title = "Land temperature - Global overview",
       y="LandTemp Cº",
       color="Temperature Type",
       caption = "fig. 4")+
  scale_color_manual(values= c("red","black","blue"))
       

GlobalCountryTemp %>% filter(Country %in% countries) %>% filter(year(Date)>1984) %>% autoplot(LandAvgTemp)+
  geom_line(data=GlobalTempV ,aes(Date, LandAvgTemp, col= "Global"),lwd=2)+
  labs(title= "Global & Country Land Temperature - Last 30 years",
            subtitle = "Average",
            y= "LandAvgTemp Cº",
       caption = "fig. 5")+
  ylim(-30,40)+
  theme(legend.position = "right")
```
! @Rodrigo: You should also add a small description for fig. 4. What can we see?

  Figure 5 shows the average temperature in the last 30 years of three selected countries and "global". As we start looking at these time series plots we can see that the temperature behaves differently according to the countries analysed, some countries belong to hot regions like Mali and even Portugal and therefore have higher temperatures while other countries like Greenland are in more northern zones and have colder temperatures. The global temperature is nothing else but the weighted average of the temperature of all countries in the original dataset and therefore it appears on the "middle" of the time series graph. The same reasoning can be applied to the max and min global temperature in figure 4.

  When we take a closer look to these graphs it's possible to start observing some seasonal and trending patterns.
Regarding the trending patterns, we can clearly see a very smooth crescent trend when it comes to land and ocean temperatures (fig. 1), which makes sense as a result of a direct effect of the global warming in the temperature, making it rise. On the other hand, there is a very clear decrescent trending on the temperature uncertainty (fig. 3), which makes sense from a point where as the time pass, the practices to measure temperature are better and more precise.

  Looking now at the seasonal pattern, when we highlight the two graphs starting from 1985 onwards (fig. 4 and 5), it becomes explicit that there is a seasonal behavior every year. There are four seasons (winter, spring, summer and autumn) and we can clearly see that each of those has its temperature characteristics.

### Focusing on seasonality, cycles and trends 

  In this section, the objective is to focus on the seasonal and trending pattern identified previously and confirm them. 

```{r, warning=FALSE}
#seasonal and trending
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  gg_season(LandAvgTemp)+
  labs(title = "Seasonality in global temperature - Land",
       y="LandAvgTemp Cº",
       color="Years",
       caption = "fig. 6")

GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  gg_season(LandOceanAvgTemp)+
  labs(title = "Seasonality in global temperature - Land and Ocean",
       y="LandOceanAvgTemp Cº",
       color="Years",
       caption = "fig. 7")
```


  These graphs (figure 5 and 6) give us some very interesting insights on the previous spotted seasonality.
  
  We can clearly see that in winter months (December, January and February), the temperature reaches the lowest values while in summer months (June, July, August), the temperature reaches its peak. Besides showing the seasonality, this graph clearly shows the crescent trending as the years go by. The lines that represent the years keep going up in the graph.
  
  Another very important aspect that we can spot is that in years around 1800 the values have some kind of spikes, this results from the fact that there was a big uncertainty around the temperature in these years and as time pass by, the uncertainty decreases and the lines become more stable. The land and ocean seasonality graph clearly shows that difference as it doesn't have values below 1900.


```{r, warning=FALSE}
#subseries
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  gg_subseries(LandAvgTemp)+
  labs(title = "Subseries - monthly global analysis",
       y="LandAvgTemp Cº",
       caption = "fig. 8")

GlobalCountryTemp %>% 
  filter(Country %in% countries2) %>% 
  gg_subseries(LandAvgTemp)+
  labs(title = "Subseries - monthly analysis",
         y="LandAvgTemp Cº",
       caption = "fig. 9")

GlobalCountryTemp %>% 
  filter(Country %in% countries3) %>% 
  gg_subseries(LandAvgTemp)+
  labs(title = "Subseries - monthly analysis",
         y="LandAvgTemp Cº",
       caption = "fig. 10")

```

  The subseries analysis also provides some more details regarding the seasonality and trending.
  
  Especially figure 8 shows clearly high discrepancies in early 1800 temperature values due to the high uncertainty. However, it is still clear to see that there is a growing trend, at a global perspective.
  
  The country subseries analysis in figure 9 and 10 reveals again a strong seasonality. However, it also shows a new perspective regarding the seasonality. Some countries have their seasonality "swaped" when compared to the global seasonality. This is due to the fact that the temperature in the countries of the northern hemisphere is the opposite of that in the countries of the southern hemisphere: If it's summer in the north (June, July, August), it's winter in the south. Nevertheless we can still identify the growing trend of the temperature in most countries.


### Looking at autocorrelation and white noise

Lastly, we analyse the autocorrelation and examine whether white noise exists.

```{r}
#6, 12 or 24 lags--> monthly data
#In this case 12 looks the best as it captures all months of one year

GlobalCountryTemp %>% filter(Country=="Global") %>% ACF(LandAvgTemp,lag_max = 12) %>% autoplot()+
  labs(title = "Autocorrelation in LandAvgTemp - Global",
       subtitle = "12 month lag",
       caption = "fig. 11")

GlobalCountryTemp %>% 
  filter(Country %in% countries2)%>% ACF(LandAvgTemp,lag_max = 12) %>% autoplot()+
  labs(title = "Autocorrelation in LandAvgTemp - Country",
       subtitle = "12 month lag",
       caption = "fig. 12")

#transforming it into yearly data since there is no seasonality
#showing the first 48 years
uncertainty<- GlobalCountryTemp %>% filter(Country=="Global") %>%
  mutate(Year= year(Date)) %>% 
  as_tibble(GlobalCountryTemp) %>% 
  group_by(Year)%>%
  select(-Date) %>%
  summarise(LandAvgTempUnc= mean(LandAvgTempUnc))

uncertainty %>%as_tsibble(index=Year) %>% filter(Year> "1752") %>% 
  ACF(LandAvgTempUnc, lag_max = 48) %>% autoplot()+
  labs(title = "Autocorrelation in LandAvgTempUnc - Global",
       subtitle = "48 years lag",
       caption = "fig. 13")
```



  By looking at  the autocorrelation plots, regarding the Land Average Temperature (see figure 11 and 12), we used a lag of 12 months since it's more appropriate for monthly data and it captures all the months of one year. It clearly shows that autocorrelation exists between the time series and its lags period. We can also spot that the autocorrelation peaks on month 6 and 12 which correspond to the highest and lowest temperature peeks of the seasonality, respectively.

  For the autocorrelation analysis of Land Average temperature Uncertainty (see figure 13), we first changed the monthly data to yearly data as there is no seasonality and it helps on the visualization. Moreover, we defined a lag period of 48 years which is about 1/3 of our total data. We can observe that it follows the typical trend behavior: Small lags have larger values as they are close to the previous value and tend to get smaller as the time goes by. 
  
  Since our autocorrelation values are mostly above the significant interval our time series don't follow "white noise".
  
  
  




# 3. Time Series Decomposition

First of all, we are replacing missing values, by the last available value for that variable (in a up-down manner), so that we keep our time series without missing gaps, to be able to decompose it.

```{r, warning = False}

GlobalCountryTemp %>%
  fill(LandAvgTemp, .direction = 'down') %>%
  fill(LandOceanAvgTemp, .direction = 'down') -> GlobalCountryTemp 
  
#colSums(is.na(GlobalCountryTemp))

```


Now we are going to see if the series need adjustments. As we can see in the graphs below, our variation does not change with the level of the time series. So, we will not use mathematical transformations. 


```{r, warning = False}

GlobalTemp %>%
  filter(year(Date)>1838, year(Date)<1900) %>%
  autoplot(LandAvgTemp)+
  geom_line(color = '#0088cc')+
  labs(title = "Land Temperature- Global",
       subtitle = "Average",
       y="LandAvgTemp Cº")

GlobalTemp %>%
  filter(year(Date)>1900, year(Date)<1990) %>%
  autoplot(LandAvgTemp)+
  geom_line(color = '#ff3399')+
  labs(title = "Land Temperature- Global",
       subtitle = "Average",
       y="LandAvgTemp Cº")

GlobalTemp %>%
  filter(year(Date)>1990, year(Date)<2014) %>%
  autoplot(LandAvgTemp)+
  geom_line(color = '#cc33ff')+
  labs(title = "Land Temperature- Global",
       subtitle = "Average",
       y="LandAvgTemp Cº")

```

As the variations around the trend are of similar magnitudes over time, we will use an Additive Decomposition. There is a consistent up and down wave regarding the seasonality, and the trend is going upward, also in a consistent way, although slowly over time. There is no exponential growth.



##Decomposing using Classical Decomposition

```{r, warning = False}

GlobalTemp %>%
  select(LandOceanAvgTemp) -> land_ocean_global_temp

land_ocean_global_temp %>%
  model(classical_decomposition(LandOceanAvgTemp, type = "additive")) %>%
  components() -> classic_decmp

classic_decmp %>%
  autoplot() +
  labs(title = "Land and Ocean Global Temperature Classical Additive Decomposition")

classic_decmp %>%
  autoplot(random, color = '#ff668c') +
  labs(title = "Variations not Captured by Classical Decomposition in Land and Ocean")

```


How is Classical Decomposition for our time series?

As Classical Decomposition assumes that the seasonal component repeats from year to year, it fits reasonably in our time series, since temperatures as always higher in the Summer and lower in the Winter, and this pattern in kept. 

Now, we will try to decompose the series into the Seasonal components, the Trend component and a Remainder, using STL Decomposition.


##Decomposing using STL

```{r, warning = False}


dec <- land_ocean_global_temp %>% 
  model(stl = STL(LandOceanAvgTemp))

#components(dec)


GlobalTemp %>%
  select(LandAvgTemp) -> land_global_temp

  
dec2 <- land_global_temp %>% 
  model(stl = STL(LandAvgTemp))

#components(dec2)

```


Isolating the trend and the seasonal adjusted temperatures in a graph.


```{r, warning = False}

land_ocean_global_temp %>%
  autoplot(LandOceanAvgTemp, color = 'gray') + 
  autolayer(components(dec), trend, color = '#6666ff') + 
  ggtitle('Land and Ocean Global Temperature Trend')


land_global_temp %>%
  autoplot(LandAvgTemp, color = 'gray') + 
  autolayer(components(dec2), trend, color = 'purple') + 
  ggtitle('Land Global Temperature Trend')

components(dec) %>%
  filter(year(Date)>1900) -> comp_after1900

land_ocean_global_temp %>%
  filter(year(Date)>1900) %>%
  autoplot(LandOceanAvgTemp, color = 'gray') + 
  autolayer(comp_after1900, trend, color = '#6666ff') + 
  ggtitle('Land and Ocean Global Temperature Trend')

land_global_temp %>%
  autoplot(LandAvgTemp, color = 'gray') + 
  autolayer(components(dec2), season_adjust, color = 'purple') + 
  ggtitle('Seasonal Adjusted Land Global Temperature')


```

In the first graph, it is possible to depict a notable growth in land and ocean temperatures trend over the years. When comparing this graph with the second one (depicting the Land Global temperatures trend), we can see that the second one's trend is not growing as significantly. This may mean that the rising trend in the oceans temperature is contributing highly to the upward trend in land plus oceans temperature.

In the third graph, zooming in to more recent years, we can see the evident growth in land and ocean global average temperatures.  

In the last graph, after adjusting the series regarding seasonal variations, we can clearly see that the land average global temperature is rising, more significantly from 1950 to the current years.




```{r, warning = False}

components(dec) %>%
  autoplot() + ggtitle('Land and Ocean Global Temperature STL Decomposition')

#components(dec2) %>%
#  autoplot() + ggtitle('Land Global Temperature') 

components(dec) %>%
  autoplot(remainder, color = '#ff668c') + ggtitle('Land Global Temperature Remainder') 

#components(dec2) %>%
#autoplot(remainder, color = '#8c66ff') + ggtitle('Land and Ocean Global Temperature Remainder') 



```

How is STL for our time series?

Looking at the remainder, there is no clear pattern. That is a sign that the STL model is capturing our season and our trend in an appropriate manner.



##Decomposing using SEATS Decomposition

```{r, warning = False}

land_ocean_global_temp %>% 
  filter(year(Date) > 1990) %>%
  model(seats = feasts::: SEATS(LandOceanAvgTemp)) %>%
  components() -> seats_dec

autoplot(seats_dec) + labs(title = "Land and Ocean Global Temperature X-13 ARIMA-SEATS Decomposition")
```

Did SEATS help?

The trend is smoother. It is easiear to see that the trend is going up.



##Trend in different Continents

```{r, message=FALSE, warning = False}

#----------------------------------------------------------------

GlobalCountryTemp %>%
  filter(year(Date)> 1850) %>%
  filter (Country == 'Europe') %>%
  select(LandAvgTemp) %>%
  as_tibble() %>%
  group_by(Date) %>% 
  summarise_at(vars(LandAvgTemp), list(LandTempEurope = mean)) %>%
  as_tsibble() %>%
  fill(LandTempEurope, .direction = 'down') -> tempEurope

decEurope <- tempEurope %>% 
  model(stl = STL(LandTempEurope))

#components(decEurope)

tempEurope %>%
  autoplot(LandTempEurope, color = 'gray') + 
  autolayer(components(decEurope), trend, color = '#9900cc') + 
  ggtitle('Land Temperature Europe Trend')

#----------------------------------------------------------------

GlobalCountryTemp %>%
  filter(year(Date)> 1860) %>%
  filter (Country == 'Africa') %>%
  select(LandAvgTemp) %>%
  as_tibble() %>%
  group_by(Date) %>% 
  summarise_at(vars(LandAvgTemp), list(LandTempAfrica = mean)) %>%
  as_tsibble() %>%
  fill(LandTempAfrica, .direction = 'down') -> tempAfrica

decAfrica <- tempAfrica %>% 
  model(stl = STL(LandTempAfrica))

#components(decAfrica)

tempAfrica %>%
  autoplot(LandTempAfrica, color = 'gray') + 
  autolayer(components(decAfrica), trend, color = '#e65c00') + 
  ggtitle('Land Temperature Africa Trend')

#----------------------------------------------------------------

GlobalCountryTemp %>%
  filter(year(Date)> 1850) %>%
  filter (Country == 'Asia') %>%
  select(LandAvgTemp) %>%
  as_tibble() %>%
  group_by(Date) %>% 
  summarise_at(vars(LandAvgTemp), list(LandTempAsia = mean)) %>%
  as_tsibble() %>%
  fill(LandTempAsia, .direction = 'down') -> tempAsia

decAsia <- tempAsia %>% 
  model(stl = STL(LandTempAsia))

#components(decAsia)

tempAsia %>%
  autoplot(LandTempAsia, color = 'gray') + 
  autolayer(components(decAsia), trend, color = '#ff4d94') + 
  ggtitle('Land Temperature Asia Trend')

#----------------------------------------------------------------

GlobalCountryTemp %>%
  filter(year(Date)> 1850) %>%
  filter (Country == 'South America') %>%
  select(LandAvgTemp) %>%
  as_tibble() %>%
  group_by(Date) %>% 
  summarise_at(vars(LandAvgTemp), list(LandTempSouthAmerica = mean)) %>%
  as_tsibble() %>%
  fill(LandTempSouthAmerica, .direction = 'down') -> tempSouthAmerica

decSouthAmerica <- tempSouthAmerica %>% 
  model(stl = STL(LandTempSouthAmerica))

#components(decSouthAmerica)

tempSouthAmerica %>%
  autoplot(LandTempSouthAmerica, color = 'gray') + 
  autolayer(components(decSouthAmerica), trend, color = '#008040') + 
  ggtitle('Land Temperature South America Trend')

#----------------------------------------------------------------

GlobalCountryTemp %>%
  filter(year(Date)> 1850) %>%
  filter (Country == 'North America') %>%
  select(LandAvgTemp) %>%
  as_tibble() %>%
  group_by(Date) %>% 
  summarise_at(vars(LandAvgTemp), list(LandTempNorthAmerica = mean)) %>%
  as_tsibble() %>%
  fill(LandTempNorthAmerica, .direction = 'down') -> tempNorthAmerica

decNorthAmerica <- tempNorthAmerica %>% 
  model(stl = STL(LandTempNorthAmerica))

#components(decNorthAmerica)

tempNorthAmerica %>%
  autoplot(LandTempNorthAmerica, color = 'gray') + 
  autolayer(components(decNorthAmerica), trend, color = '#009999') + 
  ggtitle('Land Temperature North America Trend')

#----------------------------------------------------------------

GlobalCountryTemp %>%
  filter(year(Date)> 1850) %>%
  filter (Country == 'Oceania') %>%
  select(LandAvgTemp) %>%
  as_tibble() %>%
  group_by(Date) %>% 
  summarise_at(vars(LandAvgTemp), list(LandTempOceania = mean)) %>%
  as_tsibble() %>%
  fill(LandTempOceania, .direction = 'down') -> tempOceania

decOceania <- tempOceania %>% 
  model(stl = STL(LandTempOceania))

#components(decOceania)

tempOceania %>%
  autoplot(LandTempOceania, color = 'gray') + 
  autolayer(components(decOceania), trend, color = '#99004d') + 
  ggtitle('Land Temperature Oceania Trend')

```
Analyzing trends on different continents, we can reach some conclusions:

It is important to make the distinction, that the trend over time is ranging from very different values from continent to continent. While in North America the temperatures are much colder in general, in Europe and Asia they are slight more moderate. In Oceania the temperatures are hotter, ranging always above 20Cº, but in Africa and South America the temperatures are alarmingly higher, especially in Africa where temperatures start at about 23ºC  in 1860, and have already over-passed the 25ºC.

In Africa and South America the trend is going upwards significantly, and much faster, compared to other continents.
In Asia and North America, the trend is increasing, but more slowly than in other continents.


## Trend and Seasonal Strenght in different Countries and Continents


```{r}
europe <- c('Austria', 'Portugal', 'Germany', 'United Kingdom (Europe)', 'Norway',
           'France (Europe)', 'Ukraine', 'Poland', 'Greece', 'Belgium',
            'Hungary', 'Italy', 'Albania', 'Spain')

africa <- c('Algeria', 'Angola', 'Benin', 'Botswana', 'Burkina Faso',
            'Burundi', 'Cape Verde', 'Mali', 'Congo (Democratic Republic Of #The)', 'Zambia', 'Zimbabwe', 'Morocco', 'Niger', 'Nigeria')


asia <- c('China', 'Russia', 'Malaysia', 'Japan', 'Afghanistn', 'Saudi Arabia', 
          'Taiwan', 'Qatar', 'Armenia', 'Azerbaijan', 'Bahrain',
          'Bangladesh', 'Bhutan', 'Brunei', 'Cambodia', 'Cyprus', 
          'Georgia', 'India', 'Indonesia', 'Iran', 'Jordan', 'Yemen')

north_america <- c('Antigua And Barbuda', 'Bahamas', 'Barbados', 
                   'Belize', 'Canada', 'Costa Rica', 'Cuba', 'Dominica',
                   'Dominican Republic', 'El Salvador', 'Grenada', 
                   'Guatemala', 'Haiti', 'Honduras', 'Jamaica', 'Mexico',
                   'United States', 'Panama')

south_america <- c('Argentina', 'Bolivia', 'Brazil', 'Chile',
                   'Colombia', 'Peru', 'Paraguay', 'Guyana', 'Ecuador',
                   'Suriname', 'Uruguay', 'Venezuela')

oceania <- c('Australia', 'Fiji', 'Kiribati', 'Marshall Islands', 
             'Micronesia', 'Nauru', 'New Zealand', 'Palau', 'Papua New Guinea',
             'Samoa', 'Tonga')


GlobalCountryTemp %>% 
  fill(LandAvgTemp, .direction = "down") %>%
  features(LandAvgTemp, feat_stl) %>%
  drop_na(trend_strength, seasonal_strength_year) %>%
  mutate(Continent = case_when(
    Country %in% europe  ~ "Europe" ,
    Country %in% africa  ~ "Africa" ,
    Country %in% asia  ~ "Asia" ,
    Country %in% north_america  ~ "North America",
    Country %in% south_america  ~ "South America",
    Country %in% oceania  ~ "Oceania"
  )
) %>%
  drop_na(Continent) %>%
  ggplot(aes(x = trend_strength, y = seasonal_strength_year, colour = Continent)) +
  scale_color_manual(values=c('#ff9900', '#ff4d94', '#9900cc', '#009999', '#ffcc00', '#00cc00')) +
  geom_point() +
  labs(title= 'Trend and Seasonal Strenght in different Countries')


```

In the scatter plot above, we can evaluate the strength of the trends and seasonal patterns on different countries, grouped by continent.

In all countries in the graph, the seasonal patterns are clearly strong (the majority with a seasonal strength higher than 0.7).

In Europe, we can see countries are in a cluster with a very high seasonal strength (the differences in temperature according to months, are more pronounced) and lower trend strength (it is not going on an upward trend as much). 

In Asia the seasonal strength is very high in almost all plotted countries (above 0.85). Meaning the differences in temperatures in distinct seasons are very significant (colder winters and very hot summers/spring).

In North America, the trend is very string in most of the countries, meaning temperatures are rising quite fast and significantly.




## Subseries of Land and Ocean Global Temperature

```{r, warning = False}

components(dec) %>% gg_subseries(season_year) + 
  labs(title = 'Subseries by month of Land and Ocean Global Temperatures') 

```

Moreover, on the next plot we can see that on average both the global land temperature and the global land and ocean temperature are much higher in Summer months, compared to the rest, and much lower in January, February, and December. Interestingly in July and August the trend is going down, so in previous years, temperatures were higher in these months, compared to recent years.




## 4. The forecaster’s toolbox

In the previous chapters we have already prepared and visualized our data in order to gain a good understanding of our data. Now, we will use some benchmark forecasting methods to forecast our time series. Moreover, techniques for calculating forecast intervals and methods for assessing forecast accuracy will be applied.


### Specifying a model

Looking at our data allows us to identify common patterns, and subsequently specify an appropriate model. We have seen in the previous chapters that our time series for Land Average Temperature and Land & Ocean Average temperature reveal a seasonal and a trend pattern. Now, we will check which type of forecasting method is appropriate for this type of time series.

There are several simple forecasting methods. The "Average method" and the "Naïve method" are suitable for time series without any seasonality or trend pattern. Therefore, we won't use these methods for our time series.

The "Seasonal naïve method" is appropriate for data with seasonality and the "Drift Method" for data with a trend pattern. For that reason, we will know apply these forecasting methods to our data and visualize the results.

```{r}
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  fill(LandAvgTemp, .direction = "down") %>% 
  model(Seasonal_naive = SNAIVE(LandAvgTemp),
        Drift = RW(LandAvgTemp ~ drift())
        ) %>% forecast(h = "8 years") %>% 
  autoplot((filter(GlobalCountryTemp, year(Date) > 2000))) + 
  labs(title = "Global Land Average Temperature - Forecast",
       subtitle = "Drift and Seasonal naïve method",
       y="LandAvgTemp Cº",
       caption = "fig. 20")
```


```{r}
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  fill(LandOceanAvgTemp, .direction = "down") %>% 
  model(Seasonal_naive = SNAIVE(LandOceanAvgTemp),
        Drift = RW(LandOceanAvgTemp ~ drift())
        ) %>% forecast(h = "8 years") %>% 
  autoplot(filter(GlobalCountryTemp, year(Date) > 2000)) +
  labs(title = "Global Land & Ocean Average Temperature - Forecast",
       subtitle = "Drift and Seasonal naïve method",
       y="LandOceanAvgTemp Cº",
       caption = "fig. 21")
```

The figures above show the forecasts for the next 8 years (from 2016 onward) estimated by the "Seasonal naïve method" and the "Drift Method" for the times Series "Global Land Average temperature" and "Global Land & Ocean Temperature", respectively. It can be seen that the area for the prediction intervals is high (espicially for the "Drift method"), which means that there is a high uncertainty in the forecasts. One reason for this may be that one method is suitable for data with seasonality and the other method for data with trend. But in fact our time series has both seasonality and trend. 

Let's try to apply the "Drift Method" on the seasonal adjusted data, so that we capture both components in our forecast: Seasonality and Trend.


```{r, warning = FALSE}
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  fill(LandAvgTemp, .direction = "down") %>%
  model(stlf = decomposition_model(
    STL(LandAvgTemp ~ trend(window = 7), robust = TRUE),
        Drift = RW(season_adjust ~ drift())
    )) %>% 
  forecast(h = "8 years") %>% 
  autoplot(filter(GlobalCountryTemp, year(Date) > 2000)) +
  labs(title = "Global Land Average Temperature - Forecast",
       subtitle = "Drift method based on seasonal adjusted data",
       y="LandAvgTemp Cº",
       caption = "fig. 22")

```

```{r, warning = FALSE}
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  fill(LandOceanAvgTemp, .direction = "down") %>%
  filter(!is.na(LandOceanAvgTemp)) %>% 
  model(stlf = decomposition_model(
    STL(LandOceanAvgTemp ~ trend(window = 7), robust = TRUE),
        Drift = RW(season_adjust ~ drift())
    )) %>% 
  forecast(h = "8 years") %>% 
  autoplot(filter(GlobalCountryTemp, year(Date) > 2000)) +
  labs(title = "Global Land & Ocean Average Temperature - Forecast",
       subtitle = "Drift method based on seasonal adjusted data",
       y="LandOceanAvgTemp Cº",
       caption = "fig. 23")
```

The figures above show the forecasts for the next 8 years (from 2016 onward) estimated by the "Drift Method" based on the seasonal adjusted data (based on STL Decomposition) for the times Series "Global Land Average temperature" and "Global Land & Ocean Temperature", respectively. The prediction intervals are still big, especially for the forecast of the "Global Land Average temperature". When we compare the prediction interval to those of the "Seasonal naïve method" in figure 20 and 21 they seem to be even bigger.

In the following chapter we will calculate the forecasting accuracy and check if our intuition about the best method is true.



### Accuracy & performance evaluation

In this chapter we will divide our data in training and test sets to evaluate the forecasting accuracy. For this purpose the time series cross-validation is used. In this procedure, there are a number of test sets, each consisting of a single observation. The corresponding training set consists only of observations that occurred before the observation that forms the test set. By averaging over the test series the forecasting accuracy is calculated. 

In the following we will use the one-step ahead forecast with an initial sample size of 3 and the stretch_tsibble function (=extends a growing length window with new data) to calculate the accuracy obtained via time series cross-validation. To increase the performance of the code, only data since 1990 will be considered.


#### Global land average temperature forecast

```{r, warning = FALSE}
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  filter(year(Date)>=1990) %>%
  fill(LandAvgTemp, .direction = "down") %>%
  filter(!is.na(LandAvgTemp)) %>%
  stretch_tsibble(.init = 3, .step = 1) %>% 
  model(Seasonal_naive = SNAIVE(LandAvgTemp),
        Drift = RW(LandAvgTemp ~ drift())
        ) %>% forecast(h = "8 years") %>% 
  accuracy(GlobalCountryTemp)
```

```{r, warning = FALSE}
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  filter(year(Date)>=1990) %>%
  fill(LandAvgTemp, .direction = "down") %>%
  filter(!is.na(LandAvgTemp)) %>%
  stretch_tsibble(.init = 3, .step = 1) %>%
  model(stlf = decomposition_model(
    STL(LandAvgTemp ~ trend(window = 7), robust = TRUE),
        Drift = RW(season_adjust ~ drift())
    )) %>% forecast(h = "8 years") %>% 
  accuracy(GlobalCountryTemp)

```

In the figures above we can see the accuracy of the three different models for the Average global Land Temperature. The RMSE is a good way to choose the best forecasting model. The RMSE of the "Seasonal naïve method" is the smallest, followed by the "Drift Method" on seasonal adjusted data and then the normal "Drift method".


#### Global land and ocean average temperature forecast

```{r}
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  filter(year(Date)>=1990) %>%
  fill(LandOceanAvgTemp, .direction = "down") %>%
  filter(!is.na(LandOceanAvgTemp)) %>%
  stretch_tsibble(.init = 3, .step = 1) %>% 
  model(Seasonal_naive = SNAIVE(LandOceanAvgTemp),
        Drift = RW(LandOceanAvgTemp ~ drift())
        ) %>% forecast(h = "8 years") %>% 
  accuracy(GlobalCountryTemp)
```

```{r, warning = FALSE}
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  filter(year(Date)>=1990) %>%
  fill(LandOceanAvgTemp, .direction = "down") %>%
  filter(!is.na(LandOceanAvgTemp)) %>%
  stretch_tsibble(.init = 3, .step = 1) %>%
  model(stlf = decomposition_model(
    STL(LandOceanAvgTemp ~ trend(window = 7), robust = TRUE),
        Drift = RW(season_adjust ~ drift())
    )) %>% forecast(h = "8 years") %>% 
  accuracy(GlobalCountryTemp)

```

In the figures above we can see the accuracy of the three different models for the Average global Land & Ocean Temperature. By evaluating the RMSE we can again see that the forecasting accuracy of the "Seasonal naïve method" is the best, followed by the "Drift Method" on seasonal adjusted data and then the normal "Drift method".

The reason for this may be that our time series shows a strong seasonality but only a slight trend. Therefore, the "Seasonal naïve method" already captures the most important factor. If the forecasting horizon would be bigger it can be assumed, that the performance of the "Drift Method" on seasonal adjusted data increases. Moreover, a different value for the step ahead forecast (we chose one-step ahead forecast) could lead to a different result.

#### Checking Residuals of selected method

Based on our previous analysis, we can conclude that the "Seasonal naïve method" is the most appropriate method for our time series. Let's have a look at the residuals of this method:

```{r, warning = FALSE}
GlobalCountryTemp %>% 
  filter(Country=="Global") %>% 
  filter(year(Date)>=1990) %>%
  fill(LandAvgTemp, .direction = "down") %>% 
  model(Seasonal_naive = SNAIVE(LandAvgTemp)) %>% 
  augment() %>% 
  autoplot(.resid) + labs(title = "Residuals from Seasonal naïve method", caption = "fig. 24")

GlobalCountryTemp %>% 
  filter(Country=="Global") %>%
  filter(year(Date)>=1990) %>%
  fill(LandAvgTemp, .direction = "down") %>% 
  model(Seasonal_naive = SNAIVE(LandAvgTemp)) %>% 
  augment() %>% 
  ACF(.resid) %>% 
  autoplot() + labs(title = "ACF of Residuals from Seasonal naïve method", caption = "fig. 25")
```

We have limited our time series to data from 1990 onwards, as we do not need data from 1850 onwards. In the two figures above we can see the residuals from the "Seasonal naïve method" and the ACF of the residuals from the "Seasonal naïve method". The ACF value is for some lags very low, but we can also see that there are several outliers. Therefore, we can't say that the residuals of our selected method looks like white noise.
This means that none of the Benchmark method is optimal for our time series. However, the "Seasonal naïve method" seems to perform good with regard to the forecasting accuracy.






